Exploring dimensions of biodiversity in Japanese ferns

### 2020-07-24 ms/SI.Rmd
Update SI
- move ggplot themes to functions.R
- fix variable names for canape

```{r plot-prep}
# Define ggplot2 theme for maps
map_theme <- theme(
    panel.grid.major = element_line(color = "grey95", size = 0.1),
    panel.background = element_rect(fill = "grey60"),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
  )

map_theme_light <- theme(
    panel.grid.major = element_line(color = "white", size = 0.1),
    panel.background = element_rect(fill = "grey90"),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
  )
```

@@

```{r endemism-restricted, fig.width = 7, fig.height = 8}
# PE (endemic)
# FIXME: cells with zero species should be NA, not zero PE
a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES PE (endemic)
# FIXME: cells with zero species should be NA, not zero PE
b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_rank)) +
  geom_sf(color = "transparent") +
  labs(fill = "Rank PE") +
  scale_fill_scico(palette = "lajolla") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

# RPE
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "RPE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_light +
  theme(
    axis.text.x = element_blank()
  )

# SES RPE
d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPE") +
  scale_fill_scico(
    palette = "roma",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(axis.text.y = element_blank())

@@

e <- ses_div_ferns_endemic_spatial %>%
  mutate(endem_type = replace_na(endem_type, "none")) %>%
  ggplot(aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_fill_manual(values = canape_cols, breaks = c("neo", "paleo", "mixed", "none"), labels = c("Neo", "Paleo", "Mixed", "Not endemic")) + 
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_light

>>>>

@@

```{r endemism-restricted, fig.width = 7, fig.height = 8}
# PE (endemic)
# FIXME: cells with zero species should be NA, not zero PE
a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  theme(
    axis.text.x = element_blank()
  )

# SES PE (endemic)
# FIXME: cells with zero species should be NA, not zero PE
b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_p_upper)) +
  geom_sf(color = "transparent") +
  labs(fill = "PE p-score") +
  scale_fill_scico(palette = "lajolla") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

# RPE
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "RPE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_light() +
  theme(
    axis.text.x = element_blank()
  )

# SES RPE
d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPE") +
  scale_fill_scico(
    palette = "roma",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  theme(axis.text.y = element_blank())

@@

e <- ses_div_ferns_endemic_spatial %>%
  mutate(endem_type = replace_na(endem_type, "none")) %>%
  ggplot(aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_fill_manual(values = canape_cols, breaks = c("neo", "paleo", "mixed", "none"), labels = c("Neo", "Paleo", "Mixed", "Not endemic")) + 
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_light()
